---
title: "IPsPCA"
output: pdf_document
date: "2023-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Similar to the settings for `IPsPCA_experiment1.Rmd`

```{r}
# Packages for graph objects
library(spdep)
library(igraph)

### Packages to extract incidence matrix from a general graph
library(intergraph)
library(network)

# Computations
library(CVXR)

# plot
library(ggplot2)

# ARI; compare estimated clusters with ground truth
library(aricode)


# Import methods
source("BiIPsPCA.R")


```


Import data:
```{r}
syn.locs <- read.csv("Square_sps_location_seed1.csv", header = TRUE, row.names = 1)  ##2D coordinates
syn.counts <- read.csv("Square_sps_count_seed1.csv", header = TRUE, row.names = 1)  ##Poisson counts 
```

Construct 1-NN graph (spdep object) from the locations:
```{r}
syn.coords <-  cbind(syn.locs$x, syn.locs$y)
syn.1nn <- knearneigh(syn.coords, k = 1)
```

From spdep neighborhood object to igraph graph object (Redundancy):
```{r}
syn.nb <- knn2nb(syn.1nn)
syn.g = graph_from_adj_list(syn.nb)  
syn.e = as_edgelist(syn.g) 
syn.g = graph_from_edgelist(syn.e, directed = FALSE) 
##Here,one edge can be weighted twice if node i and j are mutual closest k-neighbors.
##Due to formatting issues, we did graph from adj_list, then edgelist from graph, eventually graph from edgelist (Redundancy)
```

Visualize the 1-NN graph (igraph object):
```{r}
plot(syn.g, layout = syn.coords,vertex.size = 2, vertex.label = NA) ## we set directed = False for plotting purpose
```

```{r}
syn.inciMat = t(as.matrix(asNetwork(graph_from_edgelist(syn.e, directed = TRUE)),matrix.type="incidence"))
```


Initialize by PCA (SVD). Here we set the number of components $k = 2$.

```{r}
X0 = data.matrix(syn.counts)
X0 = scale(X0)  ### Naive normalization 
n = dim(X0)[1]
p = dim(X0)[2]

S0 = t(X0) %*% X0 /(n-1)  ### May use n - 1 since we center the X

k = 2 ## user-chosen dimension
X0_k = svd(X0)
v0 = X0_k$v[,1:k]   ## base function: svd. This seems to be slower than NumPy
u0 = X0_k$u[, 1:k]
```

#Now, we check the convergence by measuring the norm of off diagonal elements (take the squre root of sum of squares of all off-diagonal entries)

```{r}
BiIPsPCA_result = BiIPsPCA(X0, v0,syn.inciMat, tau = 0.25, max_it =50 )   
```


```{r}
plot(log(BiIPsPCA_result$off_diags), xlab = "iteration t", ylab = expression(log(norm(Sigma^"off"))))
```

