---
title: "IPsPCA"
output: pdf_document
date: "2023-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Packages for graph objects
library(spdep)
library(igraph)

### Packages to extract incidence matrix from a general graph
library(intergraph)
library(network)

# Computations
library(CVXR)

# plot
library(ggplot2)

# ARI; compare estimated clusters with ground truth
library(aricode)


# Import methods
source("IPsPCA.R")


```


Import data:
```{r}
syn.locs <- read.csv("Square_sps_location_seed1.csv", header = TRUE, row.names = 1)  ##2D coordinates
syn.counts <- read.csv("Square_sps_count_seed1.csv", header = TRUE, row.names = 1)  ##Poisson counts 
```

Construct 3-NN graph (spdep object) from the locations:
```{r}
syn.coords <-  cbind(syn.locs$x, syn.locs$y)
syn.3nn <- knearneigh(syn.coords, k = 3)
```

From spdep neighborhood object to igraph graph object (Redundancy):
```{r}
syn.nb <- knn2nb(syn.3nn)
syn.g = graph_from_adj_list(syn.nb)  
syn.e = as_edgelist(syn.g) 
syn.g = graph_from_edgelist(syn.e, directed = FALSE) 
##Here,one edge can be weighted twice if node i and j are mutual closest k-neighbors.
##Due to formatting issues, we did graph from adj_list, then edgelist from graph, eventually graph from edgelist (Redundancy)
```

Visualize the 3-NN graph (igraph object):
```{r}
plot(syn.g, layout = syn.coords,vertex.size = 2, vertex.label = NA) ## we set directed = False for plotting purpose
```

Use network object to extract incidence matrix (Currently, it will be utilized as a dense matrix for the naive implementations and cause storage and runtime problems):

```{r}
syn.inciMat = t(as.matrix(asNetwork(graph_from_edgelist(syn.e, directed = TRUE)),matrix.type="incidence"))
```


Initialize by PCA (SVD). Here we set the number of components $k = 2$. We have $n = 980$ locations, and $p = 101$ genes. 

```{r}
X0 = data.matrix(syn.counts)
X0 = scale(X0)  ### Naive normalization 

n = dim(X0)[1]
p = dim(X0)[2] ### Here, n, p are swapped (n = 101, p = 980)

S0 = t(X0) %*% X0 / n

k = 2 ## user-chosen dimension
v0 = svd(S0)$v[,1:k]   ## base function: svd. This seems to be slower than NumPy
```

Check the alignment of $U$ from PCA with respect to the 3-NN graph:
```{r}
value(mixed_norm(syn.inciMat %*% v0 , 2, 1)) 
```


### Apply the algorithm with arbitarily chosen parameters (which roughly works. Note that a large $\tau$ will cause QR factorizaiton to fail. Convergence are not guanranteed, but the algorithm does drop to a small range of values quickly)

```{r}
IPsPCA_result = IPsPCA(S0, v0,syn.inciMat, tau = 0.05, max_it =10 )   
```

We may plot the $U$ estimated by IPsPCA and PCA.The colors represent the ground truth clusters. Note that many points in IPsPCA overlap.

```{r}
par(pty = "s")
plot(IPsPCA_result$Q[,1], IPsPCA_result$Q[,2], col = as.factor(syn.locs$group), main = "IPsPCA") 
plot(v0[,1], v0[,2], col = as.factor(syn.locs$group), main = "PCA")
```

We can then plot the estimated clusters from kmeans clustering on the $U$s:

```{r}
IPsPCA_est = data.frame(x = syn.locs$x, y = syn.locs$y, cl = as.factor(kmeans(IPsPCA_result$Q, 3)$cluster))
PCA_est = data.frame(x = syn.locs$x, y = syn.locs$y, cl = as.factor(kmeans(v0, 3)$cluster)) 

qplot(x, y , data = IPsPCA_est, col = cl, main = "IPsPCA")
qplot(x, y , data = PCA_est, col = cl, main = "PCA")
```


We use ARI to compare estimations and ground truth of clusters (ARI = 1 means perfect; the smaller the worse). The terrible performance is unsuprising since this experiment aims to show a rough analysis pipeline instead of carefully choosing parameters, etc. .

```{r}
ARI(IPsPCA_est$cl, as.factor(syn.locs$group)) 
```

```{r}
ARI(PCA_est$cl, as.factor(syn.locs$group))
```
